Missing the Point\(\colon\) Non-Convergence in Iterative Imputation Procedures

### 2020-04-15 3.Thesis/1.SimulationStudy/Writeup/manuscript.tex
Rewrote intro for peer feedback research seminar.

\begin{abstract}
(\textbf{Rewrite:}) Multple imputation by chained equations (mice) is a
widely used tool to accommodate missing data. While empirical evidence
supports the validity of inferences obtained using mice, there is no
consensus on the convergence properties of the method. This paper
provides insight into non-convergence of mice algorithms.
\end{abstract}

@@

@@

Missing data problems are ubiquitous and pervasive. \textbf{or}
Everybody who works with person data is likely to run into a missing
data problem.\\
Missingness is problematic because statistical inference cannot be
performed on the whole data without employing \emph{ad hoc} solutions
(e.g., list-wise deletion), which may yield wildly invalid results
\citep{buur18}. A popular method to obtain statistically valid results
despite of missing data is `multiple imputation' \citep[MI;][]{rubin76}.
The multiple imputation method can be roughly divided into 3 steps:
First, each missing value in an incomplete dataset is `imputed' (i.e.,
filled in) several times by some algorithmic procedure. Next, the
analysis of scientific interest is performed on each of the completed
datasets. And finally, the results of these analyses are pooled into a
single aggregated statistic.

Algorithms are often used to draw imputations from the posterior
predictive distribution of the missing values. This introduces a
potential threat to the validity of the imputations: what if the
algorithm has not converged? Are the implications then to be trusted?
And can we rely on the inference obtained on the completed data?

There is no scientific consensus on the convergence properties of MI
algorithms \citep{taka17}. Some default techniques (e.g., `predictive
mean modeling') might not yield converged states at all \citep{murr18}.
Therefore, algorithmic convergence should be monitored carefully. The
current practice is visually inspecting imputation chains for signs of
non-convergence, but this may be challenging to the untrained eye
\citep[\(\S\) 6.5.2]{buur18}. Moreover, visually inspecting imputation
chains may only diagnose severely pathological cases of non-convergence.
Therefore, diagnostic evaluation of convergence would be preferred.

Iterative imputation algorithms such as mice, are MCMC methods, and
convergence of MCMC algorithms is not from a scalar to a point but from
one distribution to another. Therefore, generated values will vary even
after convergence. The aim of convergence diagnostics for MCMC methods
is thus not to establish the point at which convergence is reached, but
to diagnose severe non-convergence.

Several MCMC convergence diagnostics exist, but it is not known
whetherthese are appropriate for MICE. The application of convergence
diagnostics to MI methods has not been systematically studied
\citep{buur18}. The open questions are, e.g.: How can non-convergence be
diagnosed? How many iterations are sufficient/needed? What are the
effects of non-convergence on estimates, predictions and inference? What
is an appropriate non-convergence diagnostic? Are the default number of
iterations sufficient? The defaults of common MI software: SPSS = 10
\href{https://www.ibm.com/support/knowledgecenter/SSLVMB_24.0.0/spss/mva/syn_multiple_imputation_impute.html}{(link)},
Mplus = 100??
\href{https://pdfs.semanticscholar.org/e20e/29e008592cbfbaa567931f74cdfdb5451405.pdf?_ga=2.55354671.54033656.1584698748-527613517.1584698748}{(link)},
Stata = 10
\href{https://www.stata.com/manuals13/mi.pdf,\%20p.\%20139}{(link)},
Amelia = NA, because resampling, not convergence
\href{https://cran.r-project.org/web/packages/Amelia/Amelia.pdf}{(link)},
en MI = 30
\href{https://cran.r-project.org/web/packages/mi/mi.pdf}{(link)}.

@@

Let \(y\) denote an \(n \times p\) matrix containing the data values on
\(p\) variables for all \(n\) units in a sample. The data value of unit
\(i\) (\(i = 1, \dots, n\)) on variable \(j\) (\(j = 1, \dots, p\)) may
be either observed or missing. The collection of observed data values in
\(y\) is denoted by \(y_{obs}\); the missing part of \(y\) is referred
to as \(y_{mis}\). (\textbf{This is not only notation, but theory too:}
For each datapoint in \(y_{mis}\), we sample \(M \times T\) times
plausible values, where \(M\) is the number of imputations
(\(m = 1, \dots, M\)) and \(T\) is the number of iterations
(\(t = 1, \dots, T\)).) The collection of samples between the initial
value (at \(t=1\)) and the final imputed value (at \(t=T\)) will be
referred to as an `imputation chain'.

@@

\textbf{(Start with current guidelines, then continue with possible
quant measures of the same things from MCMC lit.)} In iterative
algorithmic procedures (Markov chain Monte Carlo methods; e.g.,
\texttt{mice} algorithms or Gibbs samplers) non-convergence may be
present as non-stationarity within chains (i.e., trending), or as slow
mixing between chains (i.e., no intermingling (\textbf{source??})).
(\textbf{add example/guidelines visual inspection here?})

The stationarity component of convergence may be evaluated with
autocorrelation \citep[\(AC\);][]{scha97, gelm13}, numeric standard
error \citep[or `MC error';][]{gewe92}, and Raftery and Lewis's
\citeyearpar{raft91} procedure to determine the effect of trending
within chains (\textbf{or the chain length required to cancel out
trending?}).

The mixing component can be assessed with the potential scale reduction
factor \(\widehat{R}\) \citep[a.k.a. `Gelman-Rubin
statistic';][]{gelm92}. With an adapted version of \(\widehat{R}\),
proposed by \citet{veht19}, the stationarity component of convergence
might be evaluated as well. This would make \(\widehat{R}\) a general
convergence diagnostic. The application of \(\widehat{R}\) to assess
stationarity has not been thoroughly investigated. Therefore, this study
employs both \(\widehat{R}\) and autocorrelation to investigate
convergence, as recommended by \citep[p.~898]{cowl96}.

Note that all of these methods evaluate the convergence of univariate
scalar summaries (e.g., chain means or variances). These convergence
diagnostics cannot diagnose convergence of multivariable statistics
(i.e., relations between scalar summaries). \citet{buur18} proposed to
implement multivariable evaluation through eigenvalue decomposition
\citep{mack03}. This method is outside of the scope of the current study
(\textbf{not anymore?}).

@@

We can interpret \(AC\)-values as a measure of stationarity. If
\(AC\)-values are close to zero, there is no dependence between
subsequent samples within imputation chains. Positive \(AC\)-values,
however, indicate recurrence. If \(\theta\)-values of subsequent
iterations are similar, trending may occur. Negative \(AC\)-values show
no threat to the stationarity component of convergence. On the contrary
even---negative \(AC\)-values indicate that \(\theta\)-values of
subsequent iterations diverge from one-another, which may increase the
variance of \(\theta\) and speed up convergence. As convergence
diagnostic, the interest is therefore in positive \(AC\)-values.
Moreover, the magnitude of \(AC\)-values may be evaluated statistically,
but that is outside of this note's scope.

Negative \(AC\)-values indicate divergence within imputation chains.
Subsequent sampled values within each imputation chain are less alike.
It may not be wise (\textbf{why? and how would we know? e.g., wider
distrubutions? didn't seem like it\ldots{}}) to terminating the
algorithm when AC is negative. \textbf{High \(AC\)-values are
implausible in MI procedures. That is, the randomness induced by the MI
algorithm effectively mitigates the risk of dependency within chains.}

@@

Based on an empirical finding \citep{lace07}, we hypothesize that
\(\widehat{R}\) will over-estimate non-convergence of MI algorithms. The
threshold of \(\widehat{R} < 1.01\) will then be too stringent for
diagnosing convergence. This over-estimation may, however, be diminished
because \(\widehat{R}\) can falsely diagnose convergence if initial
values of the algorithm are not appropriately over-dispersed
\citep[p.~437]{broo98}. In \texttt{mice}, initial values are chosen
randomly from the observed data. Therefore, we cannot be certain that
the initial values are over-dispersed. We expect this to have little
effect on the hypothesized performance of \(\widehat{R}\). \textbf{(Add
actual hypothesis.)} No hypothesis was formulated about the performance
of \(AC\) as convergence diagnostic.

@@

@@

\textbf{Add the function \texttt{mvtnorm::rmvnorm()} back in}

The complete data is `amputed' once for each simulation repetition with
function \texttt{mice::ampute()}. (\textbf{change this:} The missingness
is univariate, and the probability to be missing is the same for all
four variables, namely 20\% (\texttt{prop\ =\ 0.8,\ mech\ =\ "MCAR"}).
This leaves 20\% of the rows completely observed). \textbf{(Add
different percentages missingness??)} This study only considers only an
MCAR missingness mechanism. Therefore, results may not be extrapolated
to other missing data problems. Proper performance of the convergence
diagnostics under MCAR is necessary but not sufficient to demonstrate
appropriateness of \(\widehat{R}\) and \(AC\) as convergence
diagnostics. This is just a proof of concept.

@@

\hypertarget{univariate-estimates-and-convergence-diagnostics-theta-mean-and-theta-variance}{%
\subsection{Univariate estimates and convergence diagnostics (theta =
mean and theta =
variance)}\label{univariate-estimates-and-convergence-diagnostics-theta-mean-and-theta-variance}}

@@

There is a clear bias in the regression estimates in simulation
conditions where the number of iterations is smaller than four. In
simulation conditions where \(T > 5\) there is little to no bias in the
estimated regression coefficients. In most simulation conditions nominal
coverage is obtained (i.e., coverage rates of .95) for the confidence
intervals around the regression coefficients. Conditions with only one
or two iterations show some undercoverage. Since confidence interval
width is stable across conditions, the undercoverage may be attributed
to the bias in the estimated regression coefficients.

@@

@@

\begin{flushleft}\includegraphics{manuscript_files/figure-latex/unnamed-chunk-5-1} \end{flushleft}

\begin{flushleft}\includegraphics{manuscript_files/figure-latex/unnamed-chunk-6-1} \end{flushleft}

\hypertarget{results-old-just-for-reference}{%
\section{Results {[}OLD, JUST FOR
REFERENCE!{]}}\label{results-old-just-for-reference}}

\hypertarget{convergence-diagnostics-1}{%
\subsection{Convergence Diagnostics}\label{convergence-diagnostics-1}}

\textbf{(Pair convergnce diagnostics with respective estimates, not
split between convergence and simulation diagnostics)}

It is apparent that there is a relation between the number of iterations
per simulation condition (\(T\)) and the convergence diagnostics.
Generally speaking, conditions with longer imputation chains (higher
\(T\)) coincide with less signs of non-convergence
(\(\widehat{R}\)-values approach one, and \(AC\)-values approach zero).

Figure 2A shows that \(\widehat{R}\)-values generally decrease with
increasing imputation chain lengths. The decline stabilizes somewhere
between the simulation conditions \(T=30\) and \(T=50\). The downward
trend is most pronounced for \(T=3\), and between \(T = 5\) and
\(T = 10\). In the intervening conditions (\(3 \leq T \leq 5\)),
however, we observe a steep increase in \(\widehat{R}\)-values. This
increase implies that non-convergence is under-estimated, and
convergence should not be diagnosed at this point (\textbf{is that
true???}). Based on the conventional threshold \(\widehat{R} < 1.2\), we
would falsely diagnose convergence at \(T=3\). According to the widely
used threshold \(\widehat{R} < 1.1\), convergence would be diagnosed for
conditions where \(T>9\). If we use the recently recommended threshold
\(\widehat{R} < 1.01\), we would conclude that convergence is reached in
none of the simulation conditions.

The \(AC\)-values displayed in Figure 2B are almost uniformly increasing
as a function of \(T\). The only \(AC\)-value that deviates from this
observation is for \(T=2\). The lowest \(AC\)-value is obtained for
\(T=3\). \%At \(T=5\), the \(AC\)-value reaches the level observed at
\(T=2\). The gradual increase plateaus between \(T=10\) and \(T=30\).
\(AC\)-values in conditions where \(T>70\) are indifferentiable from
zero, indicating stationarity. We see an initial decrease between
\(T=2\) and \(T=3\). Simulation conditions where \(T>5\) have
\(AC\)-values greater than \(T=2\) (\textbf{weird wording}).

According to this diagnostic, none of the simulation conditions show
signs of non-convergence, since we only observe negative or zero
\(AC\)-values.

Taken together, we see that \(T>3\) is the minimal requirement to
diagnose convergence (\(\widehat{R} < 1.2; AC \leq 0\)). This threshold
is, however, not sufficient, since we overlook the increase in
\(\widehat{R}\)-values up-to conditions where \(T>5\), and the
convergence diagnostics only reach stability at \(T>20\).

\hypertarget{simulation-diagnostics}{%
\subsection{Simulation Diagnostics}\label{simulation-diagnostics}}

We use average bias, average confidence interval width, and coverage
rate as performance measures to evaluate \(\widehat{R}\) and \(AC\)
(\textbf{not the goal! we want to show the effects of non-conv, not just
see how to diagnose it.}). We make the general observation that the
simulation diagnostics behave as theorized for most simulation
conditions (bias around zero, stable confidence interval widths, nominal
coverage rate at 95\%).

Figure 3A shows that bias is fairly stable across simulation conditions.
The condition \(T=1\) clearly deviates from this trend with a negative
bias. The bias for \(T=2\) is below average, but within the range of
fluctuations. Conditions where \(T>3\) can be diagnosed as unbiased, but
the average bias across iterations (as shown by a flat Loess line) only
reaches stability at \(T=20\) (\textbf{Loess might not be relevant,
check and remove accordingly!}).

CIW is only clearly divergent in the simulation condition \(T=1\), see
Figure 3B. Only conditions where \(T>3\) are therefore considered
sufficient. (\textbf{Possibly irrelevant Loess:} These conditions have
similar CIWs, but across iterations stability is not established (i.e.,
the Loess line is never completely flat within the 100 simulation
conditions).)

The empirical coverage rate across repetitions seems more or less stable
for conditions where \(T>1\), see Figure 3C. We see some over-coverage
at \(T=2\), but that is better than under-estimating the variance of
\(\bar{Q}\). On average, the coverage rate is somewhat higher than the
expected nominal coverage of 95\% \citep{neym34} (\textbf{not
anymore??}). (\textbf{Loess:} Similar to the CIWs, there is some
trending across iterations (i.e., the Loess line is never flat)).

From the simulation diagnostics, we observe that unbiased estimates with
nominal coverage rates were obtained in conditions where \(T>3\). This
suggests that as little as four iterations may be sufficient for the MI
algorithm under the current circumstances.

In short, there is a discrepancy between what the convergence
diagnostics and what the performance measures indicate. While \(T=4\)
seems sufficient with respect to simulation quantities, the condition
where \(T=4\) resulted in the `one-but-worst' values of both convergence
diagnostics. Complete algorithmic convergence as indicated by
\(\widehat{R}\) and autocorrelation is not reached in conditions where
\(T<20\).

@@

This note shows that convergence diagnostics \(\widehat{R}\) and \(AC\)
may diagnose convergence of multiple imputation algorithms, but their
performance differs from conventional applications to iterative
algorithmic procedures. (\textbf{nope! it shows that MICE can lead to
correct outcomes when they have not converged accroding to two common
conv diags. This may be due to the measures (e.g., assumption of
overdisp) or due to the Qs (lm reg coeff, not higher dimensional/more
complex RQs). Add what \%miss has to do with it.})\\
\(\widehat{R}\) and autocorrelation indicate that algorithmic
convergence may only be reached after twenty or even forty iterations,
while unbiased, confidence valid estimates estimates may be obtained
with as little as four iterations. These results are in agreement with
the simulation hypothesis: \(\widehat{R}\) over-estimates the severity
of non-convergence when applied to MI procedures. \%This may be due to
the quantity of scientific interest chosen. More `complicated' \(Q\)s
(e.g., higher order effects or variance components) might show bias,
under- or over-coverage at higher \(T\).

>>>>

\begin{abstract}
(\textbf{Rewrite:}) Multiple imputation by chained equations (mice) is a
widely used tool to accommodate missing data. While empirical evidence
supports the validity of inferences obtained using mice, there is no
consensus on the convergence properties of the method. This paper
provides insight into non-convergence of mice algorithms.
\end{abstract}

@@

\textbf{Note for peer review: things written in bold are reminders for
myself or things that I am unsure off still. Also, I just decided to
change things majorly in my results section. So right now, the figures
do not correspond to the text. Please only skim that section, because I
am definitely going to rewrite it. Thanks already! - Hanne}

@@

Missing data problems are ubiquitous and pervasive in the social and
behavioral sciences. If a dataset contains just one missing value for a
variable of interest, statistical inferences are undefined and will not
produce any result. Incomplete observations are therefore ignored by
default in many statistical packages (i.e., list-wise deletion is
employed). Unfortunately, this \emph{ad hoc} solution may yield wildly
invalid results \citep{buur18}. An alternative is to `impute' (i.e.,
fill in) every missing value in an incomplete dataset. With imputation
techniques, one or several completed datasets are created, on which
statistical inferences can be performed. The case with several completed
datasets is known as `multiple imputation' \citep[MI;][]{rubin76}, and
requires an additional step to pool the results across imputations. By
applying Rubin's rules, the pooled result will reflect the uncertainty
in the data due to missingness. Under many circumstances, this yields an
unbiased and confidence valid estimate of the inference on the true--but
missing--data \citep{buur18}. A popular method to obtain imputations is
to use the `Multiple Imputation by Chained Equations' algorithm,
shorthand `MICE'\citep{mice}. MICE is an iterative algorithmic procedure
to draw imputations from the posterior predictive distribution of the
missing values. This introduces a potential threat to the validity of
the imputations: what if the algorithm has not converged? Are the
implications then to be trusted? And can we rely on the inference
obtained on the completed data? These are all open questions, because
the convergence properties of iterative imputation algorithms have not
been systematically studied \citep{buur18}. Moreover, there is no
scientific consensus on the convergence properties of MI algorithms
\citep{taka17}. Some default MICE techniques (e.g., `predictive mean
modeling') might not yield converged states at all \citep{murr18}.
Therefore, algorithmic convergence should be monitored carefully.

The recommended practice to monitor convergence is to visually inspect
imputations for signs of non-convergence. This practice is insufficient
on two counts: 1) it may be challenging to the untrained eye, and 2)
only severely pathological cases of non-convergence may be diagnosed
\citep[\(\S\) 6.5.2]{buur18}. Therefore, a quantitative, diagnostic
evaluation of convergence would be preferred.

Iterative imputation algorithms such as MICE are Markov chain Monte
Carlo (MCMC) methods, which means that convergence is not from a scalar
to a point but from one distribution to another. The values generated by
the algorithm (e.g., imputed values) will vary even after convergence.
Therefore, the aim of convergence diagnostics for MCMC methods should
not be to establish the point at which convergence is reached, but to
monitor signs of non-convergence. Several convergence diagnostics for
MCMC methods exist, but it is not known whether these are appropriate
for MICE.

Main question: How can non-convergence be diagnosed? Sub-questions: Are
common MCMC non-convergence diagnostics appropriate for MICE? And if so,
which threshold should be used? How many iterations are
sufficient/needed to reach convergence? Are the default number of
iterations sufficient (i.e., 5 in mice, 10 in SPSS and Stata, 30 in mi)?
What are the effects of continued iterations on estimates, predictions
and inferences? And finally, do the answers differ with varying
missingness proportions?

@@

Let \(y\) denote an \(n \times p\) matrix containing the data values on
\(p\) variables for all \(n\) units in a sample. The data value of unit
\(i\) (\(i = 1, \dots, n\)) on variable \(j\) (\(j = 1, \dots, p\)) may
be either observed or missing. The collection of observed data values in
\(y\) is denoted by \(y_{obs}\); the missing part of \(y\) is referred
to as \(y_{mis}\). For each datapoint in \(y_{mis}\), we sample
\(M \times T\) times plausible values, where \(M\) is the number of
imputations (\(m = 1, \dots, M\)) and \(T\) is the number of iterations
(\(t = 1, \dots, T\)). The collection of samples between the initial
value (at \(t=1\)) and the final imputed value (at \(t=T\)) will be
referred to as an `imputation chain'. \textbf{Add: missingness
percentage = proportion of cases with one or more missing values }

@@

The two challenges for convergence of iterative algorithms are mixing
and stationarity \citep{gelm13}, see Figure 1. Without mixing,
imputation chains do not intermingle nicely. Without stationarity, there
is trending within imputation chains. While the current practice is to
inspect traceplots for signs of non-mixing and non-stationarity,
diagnostic evaluation of these two challenges would be preferred.
Non-stationarity within chains may be diagnosed with e.g.,
autocorrelation \citep[\(AC\);][]{scha97, gelm13}, numeric standard
error \citep[`MC error';][]{gewe92}, or Raftery and Lewis's
\citeyearpar{raft91} procedure to determine the effect of trending on
the precision of estimates. A widely used diagnostic to monitor mixing
between chains is the potential scale reduction factor \(\widehat{R}\)
\citep[`Gelman-Rubin statistic';][]{gelm92}. With a recently proposed
adaptation, \(\widehat{R}\) might also serve to diagnose
non-stationarity, but this has not yet been thoroughly investigated
\citep{veht19}. Therefore, mixing and stationarity will be evaluated
separately in this study. As recommended \citep[e.g.,][p.~898]{cowl96},
\(AC\) and \(\widehat{R}\) will be used.

@@

We can interpret \(AC\)-values as a measure of stationarity. If
\(AC\)-values are close to zero, there is no dependence between
subsequent samples within imputation chains.

Positive \(AC\)-values indicate recurrence. If \(\theta\)-values of
subsequent iterations are similar, trending may occur. Negative
\(AC\)-values show no threat to the stationarity component of
convergence. On the contrary even---negative \(AC\)-values indicate that
\(\theta\)-values of subsequent iterations diverge from one-another,
which may increase the variance of \(\theta\) and speed up convergence.
As convergence diagnostic, the interest is therefore in positive
\(AC\)-values. Moreover, the magnitude of \(AC\)-values may be evaluated
statistically, but that is outside of this note's scope.

Negative \(AC\)-values indicate divergence within imputation chains.
\textbf{Subsequent sampled values within each imputation chain are less
alike. High \(AC\)-values are implausible in MI procedures. That is, the
randomness induced by the MI algorithm effectively mitigates the risk of
dependency within chains.}

@@

Based on an empirical finding \citep{lace07}, we hypothesize that
\(\widehat{R}\) will over-estimate non-convergence of MI algorithms. The
threshold of \(\widehat{R} < 1.01\) will then be too stringent for
diagnosing convergence. This over-estimation may, however, be diminished
because \(\widehat{R}\) can falsely diagnose convergence if initial
values of the algorithm are not appropriately over-dispersed
\citep[p.~437]{broo98}. In \texttt{mice}, initial values are chosen
randomly from the observed data. Therefore, we cannot be certain that
the initial values are over-dispersed. We expect this to have little
effect on the hypothesized performance of \(\widehat{R}\). \textbf{(Add
actual hypothesis: we thought that mice would converge sooner than
\(\widehat{R}\) would indicate that it did)} No hypothesis was
formulated about the performance of \(AC\) as convergence diagnostic.

@@

\begin{verbatim}
# pseudo-code of simulation 
simulate data 
for (number of simulation runs from 1 to 1000)
  for (number of iterations from 1 to 100)
    create missingness
    impute the missingness
    compute convergence diagnostics
    perform analysis
    pool results
    compute simulation diagnostics
aggregate convergence and simulation diagnostics
\end{verbatim}

\textbf{(Add step for different missingness proportions.)}

@@

\textbf{(Add the function \texttt{mvtnorm::rmvnorm()} back in somewhere
here? If yes, use citation(``mvtnorm'').)}

The complete data is `amputed' once for each simulation repetition with
function \texttt{mice::ampute()}. (\textbf{change this to reflect
current simulation set-up with 5, 25, 50, 75, and 95\% of cases having
missing data:} The missingness is univariate, and the probability to be
missing is the same for all four variables, namely 20\%
(\texttt{prop\ =\ 0.8,\ mech\ =\ "MCAR"}). This leaves 20\% of the rows
completely observed). This study only considers only an MCAR missingness
mechanism. Therefore, results may not be extrapolated to other missing
data problems. Proper performance of the convergence diagnostics under
MCAR is necessary but not sufficient to demonstrate appropriateness of
\(\widehat{R}\) and \(AC\) as convergence diagnostics. This is just a
proof of concept.

@@

\begin{flushleft}\includegraphics{manuscript_files/figure-latex/unnamed-chunk-3-2} \end{flushleft}

\begin{flushleft}\includegraphics{manuscript_files/figure-latex/unnamed-chunk-3-3} \end{flushleft}

\hypertarget{univariate-estimates-and-convergence-diagnostics}{%
\subsection{Univariate estimates and convergence
diagnostics}\label{univariate-estimates-and-convergence-diagnostics}}

@@

There is a clear bias in the regression estimates in simulation
conditions where the number of iterations is smaller than four. In
simulation conditions where \(T > 5\) there is little to no bias in the
estimated regression coefficients. In most simulation conditions nominal
coverage is obtained (i.e., coverage rates of .95) for the confidence
intervals around the regression coefficients. Conditions with only one
or two iterations show some under-coverage. Since confidence interval
width is stable across conditions, the under-coverage may be attributed
to the bias in the estimated regression coefficients.

@@

\begin{flushleft}\includegraphics{manuscript_files/figure-latex/unnamed-chunk-4-2} \end{flushleft}

\begin{flushleft}\includegraphics{manuscript_files/figure-latex/unnamed-chunk-4-3} \end{flushleft}

@@

--\textgreater{}

@@

This note shows that convergence diagnostics \(\widehat{R}\) and \(AC\)
may diagnose convergence of multiple imputation algorithms, but their
performance differs from conventional applications to iterative
algorithmic procedures. (\textbf{nope! it shows that MICE can lead to
correct outcomes when they have not converged according to two common
conv diags. This may be due to the measures (e.g., assumption of
overdisp) or due to the Qs (lm reg coeff, not higher dimensional/more
complex RQs). Add what \%miss has to do with it.})\\
\(\widehat{R}\) and autocorrelation indicate that algorithmic
convergence may only be reached after twenty or even forty iterations,
while unbiased, confidence valid estimates estimates may be obtained
with as little as four iterations. These results are in agreement with
the simulation hypothesis: \(\widehat{R}\) over-estimates the severity
of non-convergence when applied to MI procedures. \%This may be due to
the quantity of scientific interest chosen. More `complicated' \(Q\)s
(e.g., higher order effects or variance components) might show bias,
under- or over-coverage at higher \(T\).
